Verification of a TCP/IP stack

### 2021-04-13 paper/paper.tex
Suite du papier

\lstdefinestyle{Spark}{
    language = Ada,
    classoffset = 1,
    morekeywords = {Socket_Struct, Socket, Not_Null_Socket, Socket_Type, Socket_Prot, IpAddr, Error_T, Port},
    classoffset = 2,
    morekeywords = {Socket_Open, Socket_Connect, Socket_Send},
    classoffset = 7,
    morekeywords = {Pre, Post},
    keywordstyle = \bfseries,
    classoffset = 0,
}

@@

    The most significant problem found in the TCP implementation relies on the concurrency.
    Concurrency is one of the biggest source of bug when writing code. It is hard to understand
    all the possible interactions between different threads which leads to many bugs that can be
    difficult to reproduce and to debug. The use of SPARK helps us to find one.

    %% TODO : Show the relevant code snippet here

@@

%% Ici on peut parler d'une ou deux typos trouvées dans le code. Mais plutôt liées à une relecture de code et pas vraiment dues à SPARK

@@

\subsubsection{Tests to validate the new implementation}

    \begin{itemize}
        \item Bug de concurrence -> SPARK
        \item Memory leak -> SPARK
        \item Problèmes trouvés.
        \item Résultats sur la nouvelle implémentation (performance, taille du code).
        \item Activation des préconditions pour voir si elles sont toutes valides.
    \end{itemize}

>>>>

\lstdefinestyle{Spark}{
    language = Ada,
    classoffset = 1,
    morekeywords = {Socket_Struct, Socket, Not_Null_Socket, Socket_Type, Socket_Prot, IpAddr, Error_T, Port},
    classoffset = 2,
    morekeywords = {Socket_Open, Socket_Connect, Socket_Send, Tcp_Send, Tcp_Wait_For_Events, Tcp_Send_Segment, Tcp_Change_State},
    classoffset = 7,
    morekeywords = {Pre, Post},
    keywordstyle = \bfseries,
    classoffset = 0,
}

@@

    The most significant problem found in the TCP implementation relies on the concurrency.
    Concurrency is one of the biggest source of bugs when writing code. It is hard to understand
    all the possible interactions between different threads which leads to many bugs that can be
    difficult to reproduce and to debug. The use of SPARK helps us to find one.
\begin{figure}[t]
\begin{lstlisting}[style=Spark, frame=bottomline, basicstyle=\ttfamily\footnotesize,
                    numbers=left, numberstyle=\tiny, escapechar=\%]
case Sock.State is
  when TCP_STATE_SYN_RECEIVED
      | TCP_STATE_ESTABLISHED =>
  -- Flush the send buffer
  Tcp_Send (Sock, Buf, Ignore_Written,      % \label{bugProg:tcpSend} %
            SOCKET_FLAG_NO_DELAY, Error);
  if Error /= NO_ERROR then return; end if;
  -- Make sure all the data has been sent out
  Tcp_Wait_For_Events   % \label{bugProg:tcpWaitForEvents} %
    (Sock       => Sock,
    Event_Mask => SOCKET_EVENT_TX_DONE,
    Timeout    => Sock.S_Timeout,
    Event      => Event);
  if Event /= SOCKET_EVENT_TX_DONE then
    Error := ERROR_TIMEOUT; return;
  end if;
  -- Send a FIN segment
  Tcp_Send_Segment      % \label{bugProg:tcpWaitForEventsPost} %
    (Sock         => Sock,
    Flags        => TCP_FLAG_FIN or TCP_FLAG_ACK,
    Seq_Num      => Sock.sndNxt,
    Ack_Num      => Sock.rcvNxt,
    Length       => 0,
    Add_To_Queue => True,
    Error        => Error);
  if Error /= NO_ERROR then return; end if;
  -- Switch to the FIN-WAIT-1 state
  Tcp_Change_State (Sock, TCP_STATE_FIN_WAIT_1); % \label{bugProg:tcpChangeState} %
\end{lstlisting}
\caption{Reproduction of the code where the concurrency bug has been found.}
\label{fig:concurrencyBug}
\end{figure}

    A snippet in figure~\ref{fig:concurrencyBug} shows a bug of this kind found in the initial code.
    At line~\ref{bugProg:tcpSend}, the buffer is flushed, if no error happened during this stage,
    the state of the socket is either \state{ESTABLISHED} or \state{CLOSE\_WAIT} because the mutex
    that protected the socket has been released and a transition has possibly be done. The call to
    \Spark{Tcp_Wait_For_Events} at line~\ref{bugProg:tcpWaitForEvents}, produces, if succeeds, the
    state of the socket to be in the set \state{ESTABLISHED}, \state{CLOSE\_WAIT} or \state{CLOSED}.
    After, those two calls, \Spark{Tcp_Send_Segment} is called to initiate a closing of the connection
    by sending a \flag{FIN} flag at line~\ref{bugProg:tcpWaitForEventsPost}.

    The programmer who wrote this code forgot that the state of the socket has changed during the last
    called and when the \flag{FIN} flag is sent, he though that the state was still \state{SYN\_RECEIVED}
    or \state{ESTABLISHED}. The transition that looks the most appropriate for him is to jump in the state
    \state{FIN\_WAIT\_1}, as done at line~\ref{bugProg:tcpChangeState}. Unfortunately, this change of
    state makes the transitions \state{CLOSE\_WAIT} $\rightarrow$ \state{FIN\_WAIT\_1} and
    \state{CLOSED} $\rightarrow$ \state{FIN\_WAIT\_1} possible. Those two transitions do not
    appear in the state machine in figure~\ref{Fig:statemachine} because they are forbidden by the
    norm.

    As the precondition given to the function \Spark{Tcp_Change_State} lists all the authorized transition,
    SPARK warmed up that an incorrect transition was possible in the current code. The code has been reported to
    the maintainer of the TCP/IP stack and fixed.

@@

%% Ici on peut parler d'une ou deux typos trouvées dans le code. Mais plutôt liées à une relecture de code et pas vraiment dues à SPARK
\subsubsection{Tests to validate the new implementation}

    In addition to the guarantees given by SPARK, we have carried out some tests by enabling the assertions at the runtime.
    Doing so strengthen our confidence in the assertions written in SPARK, and in particular it provides more sureness
    that the assertions written are not broken by C code. The assertion are checked during the execution time
    and if one of them is violated, a runtime error is raised and it is possible to investigate.

    In order to do that, we have written some HTTP drivers with different scenarios. During the different runs,
    no assertion has been violated.

@@

    \textcolor{red}{TODO : Quoi mettre ?}
